---
title: "Demo"
author: "Zian Wang"
date: "2021/12/1"
output: html_document
---
```{r, warning=FALSE}
library('foreign') ## for loading dta files using read.dta
library('ggplot2')
library("plyr") # for recoding data
library("stats")
library("cluster") ## for h-clustring
library("dplyr") ## for making the data frames
library("knitr") ## for make the table in 2-4
library("RTextTools")
library("stringdist")

theme_set(theme_bw()) # change the default ggplot theme to black-and-white

knitr::opts_chunk$set(
  echo=T, ## show your R code chunk
  message = F, ## hide the message
  warning = F, ## hide the warning
  autodep = T ## make sure your separate code chunks can find the dependencies (from other code chunk)
)
```


Read the files:

```{r, warning=TRUE}
##setwd("A:/Data_Mining/2021-fall-dm-project-angrynerds")
Sys.setlocale(category = "LC_ALL", locale = "us")
readFile <- function(filename, num) {
  text <- c()
  labels <- c()
  con <- file(filename, "r")
  line = readLines(con, n=1)
  while(length(line) != 0) {
    
    line <- gsub("\\t", "", line)
    line <- gsub("\\n", "", line)
    if(nchar(line) >= 60 && nchar(line) <= 500) {
      text <- c(text, line)
    }
    line <- readLines(con, n=1)
  }
  close(con)
  return(text)
}

text <- c()
for(i in list.files("./data")) {
  text <- c(text, readFile(paste("./data/", i, sep = "")))
}
```

Manually set the labels.

This is not neccessary for now because we have Sonal's great code, so you do not need to run this.

```{r}

label <- rep("Not problematic", length(text))
label[1:200] <- "Erases misconduct records"
label[238] <- "Erases misconduct records"
label[250] <- "Gives officers unfair access to information"
label[266] <- "Gives officers unfair access to information"
label[257] <- "Restricts/Delays Interrogations"
label[258] <- "Restricts/Delays Interrogations"
label[261] <- "Restricts/Delays Interrogations"
label[262] <- "Restricts/Delays Interrogations"
label[267:272] <- "Limits Oversight / Discipline"
label[284] <- "Gives officers unfair access to information"
label[318] <- "Requires City Pay for Misconduct"

text_2 <- readFile("./data/Florida+PBR.txt")
```

This trunk can automatically get the labels.

```{r}
### We should not need to run this everytime.
truth <- read.csv("./data.csv", header = T)
truth <- truth[c("language", "category")]

label <- rep("Not problematic", length(text))

for (i in seq_along(text)) {
  for (j in seq_along(truth$language)){
    if (stringsim(text[i], truth$language[j],method='jw', p=0.1) > 0.8){
      label[i] = truth$category[j]
    }
  }
}
```

Handel with the highly unbalanced data.

```{r}
### Read the data stored in files. If you do not have these files, do not run.
label <- readRDS("./label.rds")
text <- readRDS("./text.rds")
set.seed(123)

sub_text <- cbind(text, as.factor(label))
colnames(sub_text) <- c("Text", "Label")
sub_text <- as.data.frame(sub_text)
sub_text <- arrange(sub_text,Label)
summary(as.factor(sub_text$Label))



over_sampling <- function(sub_text, index, times) {
  temp <- data.frame()
  i <- 0
  while(i < times) {
    temp <- rbind(temp, sub_text[min(index): max(index),])
    i = i+1
  }
  return(temp)
}

under_sampling <- function(index, times) {
  temp_label <- sample(min(index):max(index), floor(length(index) * times))
  return(sub_text[temp_label,])
}

test_data <- data.frame()

get_test <- function(index, proportion) {
  test_index <- sample(index, proportion * length(index))
  return(test_index)
}

times <- 50
proportion <- 0.2

prepare_data <- function(sub_text, times, proportion) {
  t <- 1
  test_data <- data.frame()
  problematic_data <- data.frame()
  while(t < 5) {
    label <- which(sub_text$Label==t)
    test_index <- get_test(label, proportion)
    test_data <- rbind(test_data, sub_text[test_index,])
    sub_text <- sub_text[-test_index,]
    label <- which(sub_text$Label==t)
    temp <- over_sampling(sub_text, label, times)
    print(nrow(temp))
    problematic_data <- rbind(problematic_data, temp)
    t = t+1
  }
  label_5 <- which(sub_text$Label==5)
  test_index <- get_test(label_5, proportion)
  test_data <- rbind(test_data, sub_text[test_index,])
  sub_text <- sub_text[-test_index,]
  t <- 6
  while(t <= 7) {
    label <- which(sub_text$Label==t)
    test_index <- get_test(label, proportion)
    test_data <- rbind(test_data, sub_text[test_index,])
    sub_text <- sub_text[-test_index,]
    label <- which(sub_text$Label==t)
    temp <- over_sampling(sub_text, label, times)
    problematic_data <- rbind(problematic_data, temp)
    t = t+1
  }
  non_problematic_data <- sub_text %>% filter(Label == 5)
  sub_text <- rbind(problematic_data, non_problematic_data)
  sub_text <- sub_text[sample(nrow(sub_text)),]
  sub_text <- rbind(test_data, sub_text)
  sign <- data.frame()
  sign[1,1] <- "length of test_data is"
  sign[1,2] <- nrow(test_data)
  colnames(sign) <- c("Text", "Label")
  sub_text <- rbind(sub_text, sign)
}
sub_text <- prepare_data(sub_text, times, proportion)
```

The function that can assign the indexes of training data and test data.

```{r}
### Input: The start and end indexes of the testing data, the rest data will be used to train. If mode=train, the output will be the indexes of training data, if mode=test, the output will be the indexes of test data.

assign_indexes <- function(start, end, mode){
  if(mode == "train") {
    return(c(1:(start-1), (end+1):nrow(sub_text)))
  }
  else {
    return(c(start: end))
  }
}
```

Build the matrix and the container:

```{r}
length_of_test <- as.numeric(sub_text[nrow(sub_text),"Label"])
length_of_test <- as.integer(length_of_test)
sub_text <- sub_text[-nrow(sub_text),]
doc_matrix <- create_matrix(sub_text$Text, language="english", removeNumbers=TRUE,
stemWords=TRUE, removePunctuation = TRUE, removeSparseTerms = .998)

container <- create_container(doc_matrix, sub_text$Label, trainSize=4382:nrow(sub_text), testSize=1:4381, virgin=FALSE)

container <- create_container(doc_matrix, sub_text$Label[1:nrow(sub_text)], trainSize=(length_of_test+1):nrow(sub_text), testSize=1:length_of_test, virgin=FALSE)

saveRDS(sub_text, "sub_text_final.rds")
saveRDS(container, "container_final.rds")
```

Training models

Warning: for now we cannot train all the models at once, it seems the memory will overflow. So try training one model at a time.

```{r}
SVM <- train_model(container,"SVM")
SLDA <- train_model(container,"SLDA")
BOOSTING <- train_model(container,"BOOSTING")
##doc_matrix <- readRDS("doc_matrix.rds")
##container <- readRDS("container3.rds")
BAGGING <- train_model(container,"BAGGING")
RF <- train_model(container,"RF")
NNET <- train_model(container,"NNET")
TREE <- train_model(container,"TREE")

saveRDS(SLDA, "./SLDA_FINAL.rds")
saveRDS(SVM, "./SVM2.rds")
saveRDS(BOOSTING, "./BOOSTING2.rds")
saveRDS(BAGGING, "./BAGGING2.rds")
saveRDS(container, "./container3.rds")
```

```{r}
SLDA <- readRDS("./SLDA_FINAL.rds")
SVM <- readRDS("./SVM_FINAL.rds")
BOOSTING <- readRDS("./BOOSTING_FINAL.rds")
RF <- readRDS("./RF_FINAL.rds")
NNET <- readRDS("./NNET_FINAL.rds")
TREE <- readRDS("./TREE_FINAL.rds")

SLDA_CLASSIFY <- classify_model(container, SLDA)
SVM_CLASSIFY <- classify_model(container, SVM)
BOOSTING_CLASSIFY <- classify_model(container, BOOSTING)
BAGGING_CLASSIFY <- classify_model(container, BAGGING)
RF_CLASSIFY <- classify_model(container, RF)
NNET_CLASSIFY <- classify_model(container, NNET)
TREE_CLASSIFY <- classify_model(container, TREE)



analytics <- create_analytics(container,
cbind(SVM_CLASSIFY, SLDA_CLASSIFY,
BOOSTING_CLASSIFY, BAGGING_CLASSIFY,RF_CLASSIFY, 
NNET_CLASSIFY, TREE_CLASSIFY))

analytics <- create_analytics(container,
cbind(BAGGING_CLASSIFY))

summary(analytics)
topic_summary <- analytics@label_summary
alg_summary <- analytics@algorithm_summary
ens_summary <-analytics@ensemble_summary
doc_summary <- analytics@document_summary

saveRDS(analytics, "NO_OVER_SAMPLING_ANALYTICS.rds")
```

```{r}
SVM <- readRDS("./SVM2.rds")
SLDA <- readRDS("./SLDA2.rds")
BOOSTING <- readRDS("./BOOSTING2.rds")
BAGGING <- readRDS("./BAGGING2.rds")
NNET <- readRDS("./NNET2.rds")
TREE <- readRDS("./TREE2.rds")
```

Save everything, make sure you save the right things.

```{r}
saveRDS(sub_text, "subtext.rds")
saveRDS(container, "container2.rds")

saveRDS(text, "text.rds")
saveRDS(label, "label.rds")

saveRDS(SLDA, "SLDA.rds")
saveRDS(SVM, "SVM.rds")
saveRDS(BOOSTING, "BOOSTING.rds")
saveRDS(BAGGING, "BAGGING.rds")
saveRDS(NF, "NF.rds")
saveRDS(NNET, "NNET.rds")
saveRDS(TREE, "TREE.rds")
```

```{r}
SLDA <- readRDS("./SLDA.rds")
SVM <- readRDS("./SVM.rds")
BOOSTING <- readRDS("./BOOSTING.rds")
sub_text <- readRDS("./subtext.rds")
```